---
title: "Dirty data project task 4 - halloween candy data"
output:
  html_document:
    df_print: paged
---

# Initial thoughts on raw data

## Notes after first view of data:

Things to think about when I create a cleaning script:

* each file has a different number of columns

* country &  state: NOT in  2015

* gender: NOT in 2015

* age: present in all 3 files

* whether going out for halloween: present in all 3 files

* ? how to check if the list of candy bars is in the same order in all files

  * e.g. first bar "butterfinger" is in 2015 but not the others
  
* will need to pivot_longer the candy bars once I'm sure of order and have standardised the column names

* general structure:

  * 2015: date, age, going out?, then 121 candy columns
  * 2016: date, going out?, gender, age, country, state, then 117 candy columns
  * 2017: ID, going out?, gender, age, country, state, then 114 candy columns
  
* will look into data quality once I've worked out how to get the three files into a useful structure

*  might need a look-up table or similar to try and standardise country and state names

* age in different formats in different files

* invalid ages recorded in 2015

* ???what the heck are the "degrees of separation from Beyonce" type columns in 2015


## Conclusions before I start to write cleaning files:

* Write code to sort out structure using pivot_longer first - come back and look at data quality afterwards

* Maybe the column heading are not as horrible as they look - once I pivot_longer and get rid of brackets and extra characters, might not be too bad, from looking at them they seem like they might be relatively consistent between years (though I fully expect to discover more problems later!)

* I could hard code the column names for data, age etc columns - the code is going to be very specific for each year, not something that needs to be repeatable for all years

* Probably best to have a separate script for each file - could get confusing if not

* need to add a unique id before pivot



```{r}
library(here)
library(readxl)
library(tidyverse)
library(dplyr)
```


```{r}
excel_sheets(here("01_raw_data/boing-boing-candy-2015.xlsx"))
```

# Read data
```{r}
twenty_fifteen <-
  read_excel(here("01_raw_data/boing-boing-candy-2015.xlsx"))
```

```{r}
twenty_sixteen <-
  read_excel(here("01_raw_data/boing-boing-candy-2016.xlsx"))
```

```{r}
twenty_seventeen <-
  read_excel(here("01_raw_data/boing-boing-candy-2017.xlsx"))
```



# Test - pivot_longer 2015
```{r}
twenty_fifteen_reduced <-
twenty_fifteen %>% 
  select("Timestamp":"[York Peppermint Patties]" ) %>% 
  mutate(unique_id = paste(Timestamp, `How old are you?`, `Are you going actually going trick or treating yourself?`, sep = "_" ))

names(twenty_fifteen_reduced)

twenty_fifteen_long <- 
  twenty_fifteen_reduced %>% 
pivot_longer(cols = "[Butterfinger]":"[York Peppermint Patties]",
             names_to = "candy_name",
             values_to = "values",)
```

```{r}
names(twenty_fifteen)
```

