---
title: "Dirty data project - task 4 analysis"
output: html_notebook
---

This file answers the questions set by CodeClan

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(here)
```

# 1 - Prepare the data

## 1.1 - Import the files created by my cleaning scripts
There may be some errors reading in data - I'm ignoring for the moment so I can get on with the main tasks
```{r}
candy_2015_clean <-
  read_csv(here("03_clean_data/candy_2015_clean.csv"))

candy_2016_clean <-
  read_csv(here("03_clean_data/candy_2016_clean.csv"))

candy_2017_clean <-
  read_csv(here("03_clean_data/candy_2017_clean.csv"))
```


## 1.2 - Create new versions which only contain the columns I'm interested in
```{r}
candy_2015_clean <-
  candy_2015_clean %>% 
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating)

candy_2016_clean <-
  candy_2016_clean %>% 
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating)

candy_2017_clean <-
  candy_2017_clean %>% 
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating)
```

## 1.3 - Put these together into a single table
```{r}
candy_all_years <-
bind_rows(candy_2015_clean, candy_2016_clean, candy_2017_clean)
```

## 1.4 - Check the rows for each year match the individual yearly tables
```{r}
candy_all_years %>% 
  group_by(year) %>% 
  summarise(count= n())
```

## 1.5 - Drop the separate yearly tables
```{r}
rm(candy_2015_clean, candy_2016_clean, candy_2017_clean)
```



# 2 - Analysis questions set by CodeClan

## 2.1 - What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)
```{r}
candy_all_years %>%
  filter(!is.na(rating)) %>%
  summarise(total_number_ratings = n())
  
```

## 2.2 - What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

I'm not sure if I need to do something about "NA" in the age field here - but it's getting late in the day, so I'll just carry on!
```{r}
candy_all_years %>% 
  filter(!is.na(going_trick_or_treating)) %>%
  filter(!is.na(age)) %>%
  group_by(going_trick_or_treating) %>% 
  summarise(avg_age = mean(age))
```

## 2.3 - For each of joy, despair and meh, which candy bar revived the most of these ratings?

I think I should be able to do this in one query, I may come backto this later
```{r}
# JOY ratings
candy_all_years %>% 
  filter(rating =="JOY") %>% 
  group_by(candy_name) %>% 
  summarise(total_joy_ratings = n()) %>% 
  arrange(desc(total_joy_ratings)) %>% 
  head(1)

# DESPAIR ratings
candy_all_years %>% 
  filter(rating =="DESPAIR") %>% 
  group_by(candy_name) %>% 
  summarise(total_despair_ratings = n()) %>% 
  arrange(desc(total_despair_ratings)) %>% 
  head(1)

# MEH ratings
candy_all_years %>% 
  filter(rating =="MEH") %>% 
  group_by(candy_name) %>% 
  summarise(total_meh_ratings = n()) %>% 
  arrange(desc(total_meh_ratings)) %>% 
  head(1)
```

## 2.4  - How many people rated Starburst as despair?

I'm using a unique_id I created in the cleaning process to define  "people" - it's a combination of various fields in the raw data. I've had a look at the data, and am satisfied I only have one candy_name containing the string "[sS]tarburst".

This is another one I'd like to come back to with a neater bit of code.
```{r}
starburst_despair_people <-  
candy_all_years %>% 
  filter(candy_name == "Starburst") %>% 
  filter(rating == "DESPAIR") %>%
  select(unique_id, candy_name, rating) %>% 
  group_by(unique_id, candy_name, rating) %>% 
  arrange(unique_id)
  
 starburst_despair_people %>%
   group_by(rating) %>% 
   summarise(starburst_despair_people = n())
 
  
```

## 2.5 - What was the most popular candy bar by this rating system for each gender in the dataset?
```{r}
candy_all_years <-
  candy_all_years %>%
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating) %>% 
         mutate(rating_numeric = case_when(
                            rating ==  "DESPAIR" ~ -1,
                            rating ==  "JOY" ~ 1,
                            rating ==  "MEH" ~ 0,
                            TRUE             ~ NA_real_
                            
  )
  )
```
`

