---
title: "Dirty data project - task 4 analysis"
output:
  html_document:
    df_print: paged
---

This file answers the questions set by CodeClan

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(here)
```

# 1 - Prepare the data

## 1.1 - Import the files created by my cleaning scripts

```{r}
candy_2015_clean <-
  read_csv(here("03_clean_data/candy_2015_clean.csv"))

candy_2016_clean <-
  read_csv(here("03_clean_data/candy_2016_clean.csv"))

candy_2017_clean <-
  read_csv(here("03_clean_data/candy_2017_clean.csv"))
```


## 1.2 - Create new versions which only contain the columns I'm interested in
```{r}
candy_2015_clean <-
  candy_2015_clean %>% 
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating)

candy_2016_clean <-
  candy_2016_clean %>% 
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating)

candy_2017_clean <-
  candy_2017_clean %>% 
  select(year,
         unique_id,
         age,
         gender,
         country,
         going_trick_or_treating,
         candy_name,
         rating)
```

## 1.3 - Put these together into a single table
```{r}
candy_all_years <-
bind_rows(candy_2015_clean, candy_2016_clean, candy_2017_clean)
```

## 1.4 - Check the rows for each year match the individual yearly tables in the environment
```{r}
candy_all_years %>% 
  group_by(year) %>% 
  summarise(count= format(n(), big.mark= ","))
```

## 1.5 - Drop the separate yearly tables
```{r}
rm(candy_2015_clean, candy_2016_clean, candy_2017_clean)
```



# 2 - Analysis questions set by CodeClan

## 2.1 - What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

I've already dropped NA ratings in  my cleaning process after the pivot_longer stage
```{r}
candy_all_years %>%
  filter(!is.na(rating)) %>%
  summarise(total_number_ratings = format(n(), big.mark= ","))
  
```

## 2.2 - What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

I'm not sure if I need to do something about "NA" in the age field here
```{r}
candy_all_years %>% 
  filter(!is.na(going_trick_or_treating)) %>%
  filter(!is.na(age)) %>%
  group_by(going_trick_or_treating) %>% 
  summarise(avg_age = round(mean(age), 1))
```

## 2.3 - For each of joy, despair and meh, which candy bar received the most of these ratings?


```{r}

  candy_all_years %>%
  group_by(rating, candy_name) %>%
  summarise(total_ratings = n()) %>% 
  slice_max(total_ratings) %>% 
  rename("item_with_most_ratings" = "candy_name")
  
  
```


## 2.4  - How many people rated Starburst as despair?

I'm using a unique_id I created in the cleaning process to define  a person - it's a combination of various fields in the raw data.

I don't think I really need to do this - any one person can only have one row for their views on Starburst.

This is one I'd like to come back to with a neater bit of code. Though I'm quite fond of the table name **starburst_despair_people**!
```{r}
# check if there are different variations of "Starburst" in the data:
candy_all_years %>%
  filter(str_detect(candy_name, "[sS][tT][aA][rR]")) %>% 
  group_by(candy_name) %>% 
  summarise(number_people_rated_as_despair = n())
```


```{r}
# group at person level using unique_id:
starburst_despair_people <-  
candy_all_years %>% 
  filter(candy_name == "Starburst") %>% 
  filter(rating == "DESPAIR") %>%
  select(unique_id, candy_name, rating) %>% 
  group_by(unique_id, candy_name, rating) %>% 
  arrange(unique_id)
 
# response to question: 
 starburst_despair_people %>%
   group_by(candy_name, rating) %>% 
   summarise(number_of_people = n())
 
  
```


# For the next three questions, count despair as -1, joy as +1 and meh as 0
I'm adding a column for this:
```{r}
candy_all_years <-
  candy_all_years %>%
  select(year:rating) %>% 
         mutate(rating_numeric = case_when(
                            rating ==  "DESPAIR" ~ -1,
                            rating ==  "JOY" ~ 1,
                            rating ==  "MEH" ~ 0,
                            TRUE             ~ NA_real_
                            
  )
  )
```

## 2.5 - What was the most popular candy bar by this rating system for each gender in the dataset?



Note that the 2015 raw data file did not have a gender variable.

```{r}
candy_all_years %>%
  filter(!is.na(gender)) %>% 
  group_by(gender, candy_name) %>% 
  summarise(popularity_value = sum(rating_numeric)) %>% 
  slice_max(popularity_value) %>% 
  rename("most_popular_item" = "candy_name")
  
```

## 2.6 - What was the most popular candy bar in each year?

```{r}
candy_all_years %>%
  group_by(year, candy_name) %>% 
  summarise(popularity_value = sum(rating_numeric)) %>% 
  slice_max(popularity_value) %>% 
  rename("most_popular_item" = "candy_name")
```


## 2.7 - What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
candy_all_years %>% 
  group_by(country_summary = case_when(
                                      country == "USA" ~ "1. USA",
                                      country == "Canada" ~ "2. Canada",
                                      country == "UK" ~ "3. UK",
                                      TRUE ~  "4. All other countries"
                                      ),
          candy_name) %>% 
  summarise(popularity_value = sum(rating_numeric)) %>% 
  slice_max(popularity_value) %>% 
  arrange(country_summary) %>% 
  rename("most_popular_item" = "candy_name")
```


