---
title: "Dirty data project - task 3 analysis"
output:
  html_document:
    df_print: paged
---

This file answers the questions set by CodeClan, and contains some of my own investigations of the data.

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(here)
```


# 1 -  Import csv files created by *clean_data_seabirds.R*

```{r}

bird_data_by_record_id_clean <-
read_csv(here("03_clean_data/bird_data_by_record_id_clean.csv"))

ship_data_by_record_id_clean <-
read_csv(here("03_clean_data/ship_data_by_record_id_clean.csv"))

bird_data_codes_clean <-
read_csv(here("03_clean_data/bird_data_codes_clean.csv"))

ship_data_codes_clean <-
read_csv(here("03_clean_data/ship_data_codes_clean.csv"))
```



# 2 - Analysis questions set by CodeClan

## 2.1 - Which bird had the most individual sightings?
Note that the answer, Wandering albatross *sensu lato*, is an example of a record where species has not been recorded at a definitive level (you will see further examples if you expand the *head* to 10 rows in the query below).
As described in the README file for this task, I have not attempted to clean or transform records of this type.
```{r}
bird_data_by_record_id_clean %>%
group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>% 
summarise(total_sightings = n()) %>% 
arrange(desc(total_sightings)) %>% 
head(1)
```


## 2.2 - Which bird had the highest total count?
The  Short-tailed shearwater.
The data contains some very high counts for this bird  - see section 3.2 below for further investigation.
```{r}
bird_data_by_record_id_clean %>%
group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>% 
summarise(total_count = sum(count)) %>% 
arrange(desc(total_count)) %>% 
head(1)

```

## 2.3 - Which bird had the highest total count above a latitude of -30?
The Wedge-tailed shearwater:
```{r}
bird_data_by_record_id_clean %>%
  left_join(ship_data_by_record_id_clean, by = "record_id") %>% 
  filter(lat > -30) %>% 
  group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  head(1)
  
```

## 2.4 - How many different types of birds were only ever seen in groups of 1?
Twenty-three types, but please note my data cleaning assumptions: some of these species names, such as *Frigatebird (unidentified)*, are only in groups of one because the species wasn't clearly identified by the observer, and I am not making any changes to the data to try to recode this.
```{r}
bird_data_by_record_id_clean %>% 
group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>%  
summarise(max_count = max(count)) %>% 
arrange(max_count) %>% 
filter(max_count == 1)
```


## 2.5 - How many penguins were seen? (Hint: there are many types of penguin)
158, based on a search for *penguin* anywhere in the species common name:
```{r}
bird_data_by_record_id_clean %>%
filter(str_detect(species_common_name, "[pP]enguin")) %>% 
#group_by(species_common_name, 
#         species_scientific_name,
#         species_abbreviation) %>% 
summarise(total_number_penguins = sum(count)) #%>% 
#arrange(desc(total_number_penguins)) 
```


# 3 - Other analysis
## 3.1.1 - overview by year and ship activity - get data
```{r}
data_by_year_and_ship_activity  <-
  bird_data_by_record_id_clean %>%
  left_join(ship_data_by_record_id_clean, by = "record_id")  %>%
select(record_id,
        species_common_name,
        species_scientific_name,
        species_abbreviation,
        count,
        age,
        sact,
        date,
        lat,
        long,
        seasn) %>%
        mutate(year = substring(date,1,4)) %>% 
         mutate(
           ship_activity = case_when(
             is.na(sact) ~ "Not recorded",
             sact == 1 ~ "steaming, sailing",
             sact == 2 ~ "dropping trash",
             sact == 3 ~ "trawling",
             sact == 4 ~ "oceanography",
             sact == 5 ~ "potting",
             sact == 6 ~ "line fishing",
             sact == 7 ~ "cleaning fish",
             sact == 8 ~ "stationary",
             sact == 9 ~ "flying helicopters",
             sact == 10 ~ "whaling",
             TRUE       ~ "error - please check record"
           )
         ) 
         

        
  
```


## 3.1.2 - overview by year and ship activity - analysis
I've added in codes from  **ship_data_codes_clean** to give myself a quick overview of:

* which years have data

* whether the number of observations per year is consistent

* which types of shipping activity there are in the data


Probably the most useful insights for me at this stage are that:

* Not all years have observations (earliest is 1969, latest 1990, nothing for 1970-74)

* The number of observations can vary greatly from year to year

* The last couple of years of the data seem particularly low on observations (see *steaming, sailing* records for 1988-90); makes me wonder if recording "petered out" or was replaced by something else

```{r}
data_by_year_and_ship_activity %>% 
  group_by(year, ship_activity) %>% 
  summarise(number_of_records = n()) %>%
  pivot_wider(names_from = ship_activity, values_from = number_of_records)

```


## 3.2.1  - Records with very high counts

* These records may represent people recording genuinely high numbers, but they merit further investigation - someone recording the maximum number the system can cope with makes me suspicious.

*  The table **bird_data_codes_clean** states that 99,999 (the largest number it's possible to record for *count*) should be used for sightings of over 100,000 individuals, so it's possible, but I'd want to be sure before quoting this anywhere. 

* A quick Google search tells me that the Short-tailed shearwater is, for example, Australia's most numerous seabird, and that breeding colonies can contain millions of adults, so maybe I should stop being so data-skeptic and just enjoy the wonder of it!

* Perhaps the real issue is that the system used was not capable of recording full data in some situations - I wonder if this was a technical limitation at the time, or if the person who created it assumed you'd never need to record more than 100k in the count column?

* It would be  interesting to compare these high numbers to known locations of colonies: as shown in **section 3.2.2 below**, the high counts for the Short-tailed shearwater are focused in specific areas;  if these correspond with known colonies, it would reassure me that the numbers are correct (and make me wonder what the true number of individuals was!)

```{r}
data_by_year_and_ship_activity %>% 
  select(date, species_common_name, record_id, count) %>% 
arrange(desc(count)) %>% 
head(20)
```
## 3.2.2 - Records for Short-tailed shearwater where count is 99,999 shown by year, season and location

* The fact that all observations were made in spring or summer might support the idea that these were genuine records - large numbers congregate in colonies  during the breeding season

* Searching for the coordinates on Google maps gives some further evidence to support these high counts:

  * -38.00000, 1450000 is Melbourne - there is a colony of more than a million nearby: https://www.visitmelbourne.com/Regions/Phillip-Island/Things-to-do/Nature-and-wildlife/Wildlife-viewing/Bird-life

  * -41.83333, 148.6167 is off the coast of Tasmania - Tasmania has an estimated 209 colonies with 11.4 million burrows! https://dpipwe.tas.gov.au/wildlife-management/fauna-of-tasmania/birds/complete-list-of-tasmanian-birds/short-tailed-shearwater

```{r}
data_by_year_and_ship_activity %>%
  filter(species_common_name == "Short-tailed shearwater",
         count == 99999) %>% 
  select(species_common_name,
         year,
         seasn,
         lat,
         long,
         count) %>% 
  arrange(lat,
          long)
```




## 3.3 - Number of Gannets recorded per year
Just having a look at this to see if there was anything interesting.
As with **section 3.1 above**, it tells me I need to think about the number of observations there were - 
and perhaps to be careful when summarising counts by year, or comparing years.

Dividing the total count by the number of observations doesn't give any meaningful insights (you could be comparing a sighting of one individual passing through to an observation of a large group feeding), but I was just using it as a crude starting point to think about how I could provide a data series over time

e.g. perhaps to look only at observations:

* where the count was above a certain threshold
* where at least a certain number of observations were available in the year
* for a species or location where the group size might tell us something useful (e.g. number at a breeding site over time). 
```{r}
data_by_year_and_ship_activity %>%
  filter(str_detect(species_common_name, "[gG]annet"),
         ship_activity == "steaming, sailing") %>%
  group_by(year) %>% 
  summarise(total_count = sum(count),
            number_of_observations = n(),
            avg_number_of_individuals_per_observation = (sum(count)/n())
            )
```


## 3.4 - other thoughts on how I could investigate the data

* Species name: according to the table **bird_data_codes_clean**, the data uses names based on a New Zealand checklist, so I considered using the list at the link below to check/standardise species names, or perhaps to view the bird data by family or conservation status. I decided not to, so I didn't get distracted from the main tasks, but this might be interesting to come back to later.  https://avibase.bsc-eoc.org/checklist.jsp?region=NZ

* I'd be interested to come back to this data once I've learned a bit more about data visualisation techniques: it's difficult to gain any insights from just looking at the coordinates in the data, and using a mapping tool create a heat map would help me visualise this. I could also bring in other data sources - for example, comparing the information on Gannets in section 3.3 above with some of the known locations of Gannet colonies: https://teara.govt.nz/en/map/5819/new-zealands-main-gannet-colonies


