---
title: "Dirty data project - task 3 analysis"
output:
  html_document:
    df_print: paged
---


```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(here)
```

#1 -  Import bird and ship data files created by **clean_data_seabirds.R**
```{r}

bird_data_by_record_id_clean <-
read_csv(here("03_clean_data/bird_data_by_record_id_clean.csv"))

ship_data_by_record_id_clean <-
read_csv(here("03_clean_data/ship_data_by_record_id_clean.csv"))

bird_data_codes_clean <-
read_csv(here("03_clean_data/bird_data_codes_clean.csv"))

ship_data_codes_clean <-
read_csv(here("03_clean_data/ship_data_codes_clean.csv"))
```



#2 - Analysis questions set by CodeClan

##2.1 - Which bird had the most individual sightings?
```{r}
bird_data_by_record_id_clean %>%
group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>% 
summarise(total_sightings = n()) %>% 
arrange(desc(total_sightings)) %>% 
head(1)
```


##2.2 - Which bird had the highest total count?
```{r}
bird_data_by_record_id_clean %>%
group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>% 
summarise(total_count = sum(count)) %>% 
arrange(desc(total_count)) %>% 
head(1)

```

##2.3 - Which bird had the highest total count above a latitude of -30?
```{r}
bird_data_by_record_id_clean %>%
  left_join(ship_data_by_record_id_clean, by = "record_id") %>% 
  filter(lat > -30) %>% 
  group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  head(1)
  
```

##2.4 - How many different types of birds were only ever seen in groups of 1
```{r}
bird_data_by_record_id_clean %>% 
group_by(species_common_name, 
         species_scientific_name,
         species_abbreviation) %>%  
summarise(max_count = max(count)) %>% 
arrange(max_count) %>% 
filter(max_count == 1)
```


##2.5 - How many penguins were seen? (Hint: there are many types of penguin)
```{r}
bird_data_by_record_id_clean %>%
filter(str_detect(species_common_name, "[pP]enguin")) %>% 
#group_by(species_common_name, 
#         species_scientific_name,
#         species_abbreviation) %>% 
summarise(total_number_penguins = sum(count)) #%>% 
#arrange(desc(total_number_penguins)) 
```


#3 - Other analysis
##3.1.1 - overview by year and ship activity - get data
```{r}
data_by_year_and_ship_activity  <-
  bird_data_by_record_id_clean %>%
  left_join(ship_data_by_record_id_clean, by = "record_id")  %>%
select(record_id,
        species_common_name,
         species_scientific_name,
         species_abbreviation,
         count,
         age,
         sact,
        date,
        lat,
        long) %>%
        mutate(year = substring(date,1,4)) %>% 
         mutate(
           ship_activity = case_when(
             is.na(sact) ~ "Not recorded",
             sact == 1 ~ "steaming, sailing",
             sact == 2 ~ "dropping trash",
             sact == 3 ~ "trawling",
             sact == 4 ~ "oceanography",
             sact == 5 ~ "potting",
             sact == 6 ~ "line fishing",
             sact == 7 ~ "cleaning fish",
             sact == 8 ~ "stationary",
             sact == 9 ~ "flying helicopters",
             sact == 10 ~ "whaling",
             TRUE       ~ "error - please check record"
           )
         ) 
         

        
  
```


##3.1.2 - overview by year and ship activity - view summary
```{r}
data_by_year_and_ship_activity %>% 
  group_by(year, ship_activity) %>% 
  summarise(number_of_records = n()) %>%
  pivot_wider(names_from = ship_activity, values_from = number_of_records)

```


##3.2  - Records with very high counts
These records may represent people recording genuinely high numbers with the largest number the system
can cope with, but they would merit further investigation - could perhaps exclude records 
of "99999" in the meantime:
```{r}
data_by_year_and_ship_activity %>% 
  select(date, species_common_name, record_id, count) %>% 
arrange(desc(count)) %>% 
head(20)
```


## 3.3 - Number of Gannets recorded per year
Just having a look at this to see if there was anything interesting.
Mainly tells me I need to think about how many observations there were - 
perhaps need to be careful when summarising counts by year.
```{r}
data_by_year_and_ship_activity %>%
  filter(str_detect(species_common_name, "[gG]annet"),
         ship_activity == "steaming, sailing") %>%
  group_by(year) %>% 
  summarise(total_count = sum(count),
            number_of_observations = n(),
            avg_number_of_individuals_per_observation = (sum(count)/n())
            )
```



